Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-01-15 ThesisJP/Introduction/introduction.tex
Small adjustment in summary of papers.

The estimation of the spatial covariance matrix receives special attention in this study. In previous literature on Capon beamforming the matrix inversion has always been regarded as the most computationally complex piece. In this paper we show that estimation of the sample covariance matrix is actually the most complex part when spatial and temporal smoothing is included with common parameters. It is then shown how the arithmetic complexity of this estimation process can be reduced from cubic to square. Finally, we give an in-depth analysis of the arithmetic throughput on multiple platforms. The reported throughput of 1 Mpx/s on a high-end GPU is, as far as the authors know, enough to provide real-time processing for all known sonar scan sequences. However, the number of effective FLOPS is only 5 \% of the theoretical throughput of the target GPU. This result shows how hard it is to reach the theoretical level of throughput, and that how close one can get is highly algorithm depended.

@@

\subsection{Paper V}
\textbf{Huygens on Speed: Interactive Simulation of Ultrasound Pressure Fields}\\
\textbf{J.\:P.\:\AA{}sen} and S.\:Holm\\
{\it Proc. IEEE Ultrasonics Symposium 2012 \todo{add pages}.}\\\\
The fifth paper presents an interactive simulations tool capable of simulating dense pressure fields from ultrasound arrays in real time. Simulation tools are heavily used both by researchers and probe manufactures in order to test out new algorithms and to deduce the performance of new arrays prior to manufacturing. Unfortunately, simulation tools performing dense simulations are inherently slow. The tool is based on Huygens' principle which describes diffraction caused by a slit as the super position of several point sources located in the slit opening. Thus, the simulator works by accumulating the contribution from a collection of point sources in a set of observation points. This is the core engine of the simulator and it is, in the context of this thesis, of course running on a GPU. How the source and observation points are laid out is up to the client to decide. The simulator is provided with both a paint-like interface and a Matlab interface for precise scripting of array configurations. The main contributions in this paper is the increased performance of the GPU implementation compared with a CPU and Matlab (The Ultrasim toolbox) version. Second, the paint-like interface provides an interactive and neat way of demonstrating array beamforming principle. We believe this will be of great value both for array designers and teachers.

@@

Paper1

@@

>>>>

The estimation of the spatial covariance matrix receives special attention in this study. In previous literature on Capon beamforming the matrix inversion has always been regarded as the most computationally complex piece. In this paper we show that estimation of the sample covariance matrix is actually the most complex part when spatial and temporal smoothing is included with common parameters. It is then shown how the arithmetic complexity of this estimation process can be reduced from cubic to square. Finally, we give an in-depth analysis of the arithmetic throughput on multiple platforms. The reported throughput of 1 Mpx/s on a high-end GPU is, as far as the authors know, enough to provide real-time processing for all known sonar scan sequences. However, the number of effective FLOPS is only 4 \% and 14 \% of the theoretical throughput of the target GPU for the solver and covariance estimator respectively. This result shows how hard it is to reach the theoretical level of throughput, and that how close one can get is highly algorithm depended.

@@

\subsection{Paper V}
\textbf{Huygens on Speed: Interactive Simulation of Ultrasound Pressure Fields}\\
\textbf{J.\:P.\:\AA{}sen} and S.\:Holm\\
{\it Proc. IEEE Ultrasonics Symposium 2012 \todo{add pages}.}\\\\
The fifth paper presents an interactive simulations tool capable of simulating dense pressure fields from ultrasound arrays in real time. Simulation tools are heavily used both by researchers and probe manufactures in order to test out new algorithms and to deduce the performance of new arrays prior to manufacturing. Dense simulations are unfortunately inherently slow. The presented tool is based on Huygens' principle which describes diffraction caused by a slit as the super position of several point sources located inside the opening. Thus, the simulator works by accumulating the contribution from a collection of point sources in a set of observation points. This is the core engine of the simulator, and it is, in the context of this thesis, of course running on a GPU. How the source and observation points are laid out is up to the user. The simulator is provided with both a paint-like interface for interactive drawing of arrays, and a Matlab interface for precise scripting of array configurations. The main contributions in this paper is the increased performance of the GPU implementation compared with a CPU and Matlab\footnote{The Ultrasim toolbox \todo{add citation}} version. Second, the paint-like interface provides a neat way of demonstrating array beamforming principle in real time. We believe this will be of great value both for array designers and teachers.

@@

Paper1
Solving is done by a batched Gauss Jordan solver implemented by Nvidia. We decided to focus on the sample covariance estimation, and not trying to beat Nvidia's solver. A natural next step would be to analyze the current solver, and figure out if any optimizations are possible. Since the sample covariance matrix is hermitian, a solver based on Cholesky decomposition might improve the throughput for solving.

@@

No detailed evaluation of \textit{in vivo} performance. In cardiac application it has been proven hard to transfer results obtained in simulations to \textit{in vivo}. The improvements given is minimal with respect to the amount of calculation required.
